Variance reduction for MCMC methods via martingale representations

### 2020-03-12 aap/Revision/vr-mcmc-revision-new.tex
update: bias

\PVar_x[\pi_{K,n}^N(f)]=\frac{1}{n^2}\sum_{k=K+1}^{\infty}\sum_{l=1}^{n}\PE[A^2_{n-l,k}(X^x_{N+l-1})],

@@

A_{q,k}(y)=\sum_{r=0}^q \bar{a}_{r,k}(y),\quad q\in \mathbb{N}

@@

\begin{lem} 
\label{lem:bias}
Let $f$ be a bounded Borel function and $(X^x_{p})_{p \in \nset_0}$ be a Markov chain generated by ULA with constant step size $\gamma$. Then the squared bias satisfies
\begin{equation}
\label{eq:sq_bias}
\left( \mathsf{E}[\rho(f)]-\pi(f)\right)^2 \lesssim \mathrm{osc}(f)^2 \left[\gamma + \frac{1}{n}\left(V(x) + \pi_{\gamma}(V)\right)\right]
\end{equation}
for $V(x) = 1 + \|x\|^2$, where $\lesssim$ means less up to a constant not depending on $n$ and $\gamma$.
\end{lem}

\begin{proof} We can estimate the conditional bias similarly to~\cite[Section~3]{durmus:moulines:2017}:
\begin{equation}
\label{eq:06042018a2}
\left( \mathsf{E}[\pi_{K,n}^N(f)]-\pi(f)\right)^2
=\left(\mathsf{E}[\pi_{n}^N(f)]-\pi(f)\right)^2
\le
\frac{\mathrm{osc}(f)^2}{n}\sum_{p=N+1}^{N+n}
\tvnorm{\delta_x \kerULA_\gamma^{p}-\pi(\cdot)}^2,
\end{equation}
Under assumptions {\bf (H1)} and {\bf (H2)}, the corresponding Markov chain has a unique stationary distribution $\pi_\gamma$, which is different from $\pi$. From \cite[Theorem~10]{durmus:moulines:2017} and Lemma~\ref{lem:v_ergodicity} it follows that
\begin{equation}
\label{eq:bias}
\tvnorm{\pi - \pi_\gamma} < C_1\sqrt{\gamma}, \quad \tvnorm{\delta_x\kerULA_{\gamma}^{n}-\pi_\gamma} \leq C_2\rho^{\gamma n}\left(V(x) + \pi_\gamma(V)\right)
\end{equation}
for  $V(x) = 1 + \|x\|^2,$ some $\rho\in (0,1)$ and constants $C_1,C_2>0$ independent of $\gamma$  and $n$. Combining \eqref{eq:bias} with the triangle inequality, we obtain \eqref{eq:sq_bias}.
\end{proof}

>>>>

\PVar[\pi_{K,n}^N(f)]=\frac{1}{n^2}\sum_{k=K+1}^{\infty}\sum_{l=1}^{n}\PE[A^2_{n-l,k}(X^x_{N+l-1})],

@@

A_{q,k}(y)=\sum_{r=0}^q \bar{a}_{r,k}(y),\quad q\in \mathbb{N},

@@

\begin{lem} 
\label{lem:bias}
Let $f$ be a bounded Borel function and $(X^x_{p})_{p \in \nset_0}$ be a Markov chain generated by ULA with constant step size $\gamma$. Then the squared bias satisfies
\begin{equation}
\label{eq:sq_bias}
\left( \mathsf{E}[\rho(f)]-\pi(f)\right)^2 \lesssim \|f\|^2_{\infty}\left(\gamma + \frac{\rho^{2N}}{n(1-\rho^2)}\right) \lesssim \|f\|^2_{\infty}\left(\gamma + \frac{e^{-\gamma \kappa N}}{2\gamma n}\right)
\end{equation}
for $V(x) = 1 + \|x\|^2$, $\rho$ defined in Lemma~\ref{lem:v_ergodicity}, $\kappa$ independent of $N,n,\gamma$ and $\lesssim$ means less up to a constant not depending on $N,n$ and $\gamma$.
\end{lem}

\begin{proof} Under assumptions {\bf (H1)} and {\bf (H2)}, the corresponding Markov chain has a unique stationary distribution $\pi_\gamma$, which is different from $\pi$. Thus we represent the squared bias as 
\begin{eqnarray}
\label{eq:bias_1}
\left( \mathsf{E}[\pi_{K,n}^N(f)]-\pi(f)\right)^2
&=&\left(\mathsf{E}[\pi_{n}^N(f)]-\pi(f)\right)^2
=
\frac{1}{n^2}\left[\sum\limits_{k=N+1}^{N+k}\left(f(X_k^x)-\pi(f)\right)\right]^2 \leq \\
&\leq& 
2|\pi_{\gamma}(f) - \pi(f)|^2 + \frac{2}{n}\sum\limits_{k=N+1}^{N+n}\left[\PE f(X_k^x) - \pi_{\gamma}(f)\right]^2
\end{eqnarray}
 From \cite[Theorem~10]{durmus:moulines:2017} and Lemma~\ref{lem:v_ergodicity} it follows that
\begin{equation}
\label{eq:bias_2}
\tvnorm{\pi - \pi_\gamma} < C_1\sqrt{\gamma}, \quad |\PE f(X_k^x) - \pi_{\gamma}(f)| \leq C_2 \rho^n\|f\|_{V}V(x)
\end{equation}
for  $V(x) = 1 + \|x\|^2,$ $\rho\in (0,1)$ specified in Lemma~\ref{13} and constants $C_1,C_2>0$ independent of $\gamma$  and $n$. Since $f$ is bounded, $\|f\|_V \leq \|f\|_{\infty}$. Hence, combining \eqref{eq:bias_1} and \eqref{eq:bias_2}, we obtain \eqref{eq:sq_bias}.
\end{proof}
